# A list of functions available in the `SpatialSablefishAssessment` package {-}

All of these functions have descriptions and details on expected input parameters using the standard `?` R query methods. You can also use `help(package = "SpatialSablefishAssessment")` to get a list of functions available in this package.

Most of the functions will ask for an optional input parameter called `region_key` which will add a label for regions. This is a data.frame which is shown in the following code chunk. **Note** how the column `TMB_ndx` in an index made for C++ syntax not R i.e., it starts at 0 not 1.

```{r, eval = T, echo = T}
region_key = data.frame(area = c("BS","AI","WGOA","CGOA","EGOA"), TMB_ndx = c(0:4))
region_key
```


## Validate functions {-}

- `validate_input_data_and_parameters` This should always be run before `TMB::MakeADFun`

## Accessor functions {-}

- `get_AF` get age-frequency observed and predicted values from a model
- `get_LF` get length-frequency observed and predicted values from a model
- `get_catches` get observed and predicted catches from a fitted model
- `get_fishing_mortalities` get annual fishing mortalities from a model
- `get_index` get survey index from a model
- `get_partition`  get numbers at age (units 000's) from a fitted model
- `get_SSB` get SSBS (kilo tonnes 000's) from a fitted model
- `get_tag_recovery_obs_fitted_values`
- `get_tag_release_AF` get age-frequency of tag-releases
- `get_negloglike` get negative loglikelihoods for each contribution
- `simulate_observations` this function will take a TMB object and simulate a number of observations, used to create simulated residuals.

## Plotting functions {-}

- `plot_AF` plot age frequency observation
- `plot_age_length_matrix` plot input age-length transition matrices
- `plot_catch_fit` plot the fit to catches
- `plot_fishing_mortalities`  plot fishing mortalities
- `plot_frequency_of_tag_release_and_recoveries	`
- `plot_index_fit` plot survey index
- `plot_init_nage` plot initial numbers at age
- `plot_input_catches` plot input catches
- `plot_input_observations` plot input observations
- `plot_LF` plot length frequency observation
- `plot_mean_age` plot mean age observations
- `plot_mean_length` plot mean length observations
- `plot_mean_weight` plot mean weight over time for males and females
- `plot_movement`  plot movement matrix from a model
- `plot_partition` plot numbers at age by year and sex
- `plot_recruitment` plot annual recruitment by region
- `plot_selectivities` plot selectivity curves 
- `plot_SSB` plot regional SSBs
- `plot_tag_recovery_obs` plot tag-recovery obs
- `plot_tag_recovery_fits` an alternative plotting function for tag-recovery obs
- `plot_tag_release_AF` plot the numbers at age by sex for each release event

## Estimation functions {-}

- `profile_param` Run log-likelihood profiles on an estimated parameter
- `set_up_parameters` Fix parameters, uses `fix_pars` and `set_pars_to_be_the_same` for the `TagIntegrated` model 
- `pre_optim_sanity_checks` run some checks on a TMB model of type `TagIntegrated` before optimization
- `post_optim_sanity_checks`  run some checks on a TMB model of type `TagIntegrated` after optimization
- `check_gradients` check all parameter gradients are not zero
- `Francis_reweighting` calculate stage-two weights for composition data
- `get_tmb_fixed_effects` return fixed effect parameters from a model
- `fix_pars` turn off parameters so they aren't estimated, there are many parameters that shouldn't be estimated
- `set_pars_to_be_the_same` set parameters to be estimated at the same value i.e., male female have the same selectivity parameters. or tag-reporting  to be the same among regions.
- `rmvnorm_prec` simulate from a multivariate normal distribution using the precision matrix.


## Model comparison functions {-}

Often you will run multiple models to explore alternative assumptions and hypothesis. There are a range of function available for extracting output convenient for plotting and summarizing. In order to use the following functions you need to create a list containing multiple runs e.g.

```{r, echo=T, eval=FALSE}
multi_runs <- list()
multi_runs[[1]] <- low_M_model$report()
multi_runs[[2]] <- high_M_model$report()
run_labels = c("M = 0.08", "M = 0.2")

ssb_df = get_multiple_ssbs(mle_ls = multi_runs, run_labels = run_labels)
```


- `get_multiple_ssbs` get multiple models SSBs
- `get_multiple_catch_fits` Get multiple models catch fits
- `get_multiple_nlls` get multiple models negative log-likelihood 


## Distribution functions {-}

- `ddirichmult` Dirichlet-multinomial pdf function
- `dmultinom_upd` Multinomial pdf function, which is the same as TMB's. Differs from Rs base  `dmultinom` by allowing no-integers and not rounding the x values.
- `lognormal_CI` calculate confidence intervals for the lognormal distribution
- `log_cv` calculate cv for the lognormal distribution given standard deviation
- `log_sigma` calculate standard deviation for the lognormal distribution given CV



## Parameter transformation functions {-}

- `restoresimplex` go from simplex to unit vector
- `simplex` go from unit vector to simplex
- `logit_general` logistic transformation between a specified lower and upper bound
- `logit` logistic transformation between 0, 1 similar to `qlogis`
- `invlogit_general` inverse logistic transformation between a specified lower and upper bound
- `invlogit` inverse logistic transformation between 0, 1 similar to `plogis`
- `bound_unit` transform a parameter \(X\) bound between -1 and 1 (i.e., correlation parameter) as
\[
Y = 
\begin{cases}
\sqrt{\frac{X^2}{1  - X^2}}, \ \ & X \geq 0 \\
-\sqrt{\frac{X^2}{1  - X^2}} , \ \ & X < 0
\end{cases}
\]
- `inv_bound_unit` inverse of `bound_unit`
- `gm_mean` calculate the geometric mean
- `sum_to_zero_QR` take an unconstrained vector of length `N - 1` and calculate a vector of length `N` that sums = 1.
